### 2012-05-13 3.tex
include at the chapter level

\label{turkish}

%In this chapter, I present evidence that speakers 
%glossematics
%\citet{Chomsky1965} propose

%In retrospect, it seems somewhat odd that \citeauthor{Chomsky1965} would coh
%The contrast between \emph{blick} and \emph{bnick} is what has been called a ``projection problem'' \citep{Baker1979}
%This is not a universal, but rather something English speakers must learn: Moroccan Arabic, for instance, does permit \emph{bn} onsets (e.g., \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet').
%\citet{Lees1966b,Lees1966a}

%\citet{Davidson2005,Davidson2006a} shows that this ``excrescent'' schwa is distinct both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.
%\citet{Wilson2011} show that speakers use this

@@

% motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps. \citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible. \emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy. In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities \citep[][283f.]{Anderson1974}.

%a principle which is one of the two motivations  motivated the introduction of morpheme structure constraints.

%In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon does not always require static constraints on possible underlying representations, since such distributions arise by chance and static constraints defined over natural classes contribute little to predicting the contents of the lexicon. 
%This result does not directly bear on the question of whether static restrictions have any synchronic reality.

%Yet countless static constraints are statistically supported. Under the null hypothesis advanced in Chapter \ref{MSCs}, the lexical statistics of co-occurrence restrictions are irrelevant: the language acquisition device only attends to derived restrictions and ignores static restrictions regardless of their statistical reliability. On the other hand, many linguists assume that statistical significance provides sufficient conditions for positing synchronically active static constraints. For instance, \citet{Brown2010} writes:

%\begin{quotation}
%\ldots the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
%\end{quotation}

@@

\input{3.1}
\input{3.2}
\input{3.3}

>>>>

\chapter{Static and derived phonotactic constraints in Turkish} 
\label{turkish}

%This chapter attempts to derive which phonotactic judgemetns are internalized by

\citet{Chomsky1965} argue that speakers internalize not only generalizations governing phonological alternations in their language, but also about the shapes of possible words in their language. It is certainly possible to imagine that speakers would not bother at all. \citet[][320]{Zimmer1969} claims these generalizations ``have no observable consequences in the course of the normal use of the language'' and a similar sentiment is echoed more recently by \citeauthor{PE}:

\begin{quote}
Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for. \citep[][18]{PE}
\end{quote}

This is a principled null hypothesis, but one which can be quickly rejected. \citeauthor{Chomsky1965} point to robust contrasts between equally unattested items like *\emph{blick}, which is recognized by speakers a possible word of English, and *\emph{bnick}, which is not.\footnote{Any readers who are skeptical that \citeauthor{Chomsky1965}'s intuitions generalize will find support for their claim in the wordlikeness judgement data reported in Appendix \ref{albright}.}
%As a quick validation of this claim, consider the results reported by \citet{Albright2007} and discussed in the previous chapter. English speakers asked to rate non-words on a scale from 1-7, where 7 indicates ``most like English'', favor non-words [blʌs, blɑd] (average scores 4.67 and 5.13) over [bnʌs, bnɑd] (average scores 2.06, 2.00).
This preference must be regarded as a learned behavior, since many languages permit  \emph{bn} onsets (e.g., Moroccan Arabic \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet'). It is also possible to show speakers access this knowledge not just in wordlikeness judgement tasks but also in naturalistic and quotidien tasks like segmenting spontaneous speech into words, as \citet{McQueen1998} and \citet{Mattys2001b} argue for adult and infant subjects, respectively.

\citeauthor{Chomsky1965} were not the first to take an interest in characterizing the lexicon, which was also a major concern for the scholars of the Copenhagen school. In two sophisticated studies dealing with possible and impossible onsets in Georgian, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} recognize the inherent difficulty of distinguishing between structural and accidental phonotactic gaps in the lexicon on the basis of statistical criteria alone. What distinguishes the early generativist approach to phonotactics is the mentalistic concern with what generalizations speakers have internalized. By grounding the study of phonotactics in the minds of speakers, it is possible to give a precise definition to the previously vague notions; an accidental phonotactic gap is simply one which speakers do not attend to or internalize. Yet few studies of phonotactics are explicitly concerned with what speakers have internalized, despite the increasing sophistication of the phonological representations used. In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique, a chi-square test, in an attempt to validate a generalization about Arabic root consonant co-occurrence faces originally due to \citet{Greenberg1950}. While this particular generalization was ultimately also evaluated using a wordlikeness task \citep{Frisch2004}, there is now an extensive literature in which phonotactic knowledge is inferred directly from lexical statistics, with no psycholinguistic component of any kind \citep[e.g.,][]{Mester1988,Padgett1992,Padgett1995,Berkley1994b,Berkley1994a,Berkley2000,Buckley1997,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Pozdniakov2007,Anttila2008,Coetzee2008a,Dmitrieva2008a,Dmitrieva2008b,Graff2011}. The (usually tacit) assumption of most of this work is made explicit by \citet{Brown2010} in his study of consonant cluster co-occurrence restrictions in Gitksan: any statistically significant pattern in the lexicon is one that is internalized by speakers.

\begin{quote}
\ldots{}the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
\end{quote}

In the loanword adaptation literature, it has recently been recognized that not all statistically reliable generalizations about the native lexicon take part in nativization of borrowings \citep[e.g.,][]{Ito1995a,Ito1995b,Ussishkin2003}. In this chapter, data from Turkish is used to show that a widely reported, statistically reliable phonotactic generalization (\textsc{Labial Attraction}) has no effect on wordlikeness judgements. This undermines the widespread use of statistic criteria to determine which of the vast number of the overlapping, competing phonotactic generalizations are internalized by speakers and which are ignored. Comparison of \textsc{Labial Attraction} to related constraints in Turkish suggests a novel necessary condition for identifying psychologically real phonotactic constraints: namely, speakers attend to phonotactic generalizations only when they are derived from phonological processes (including syllabification).

There are a number of precedents for

This hypothesis was anticipated by \cite[][283]{Anderson1974}.

\begin{quote}
It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained. \citep[][283]{Anderson1974}
\end{quote}

This hypothesis is not ahdered to by early research on sequence structure constraints \citep[e.g.,][]{SPR,Chomsky1965,SPE,Stanley1967}.

Work in this vein sought to use sequence structure constraints to compress the lexicon as much as possible

This work has a second interest, namely in ``compressing'' the lexicon as much as possible,

It is important to know that this wor It is important to know that this wor It is important to know that this wor It is important to know that this work

% motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps. \citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible. \emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy. In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities

This condition (henceforth, the \textsc{derived} condition) 
On a historical note, this hypothesis is more constrained than what was proposed
such constraints must be derived from phonological processes in the language in question
synchronically motivated phonological processes such as alternation or syllabification.

%\footnote{I note a potential problem with this account: \citet{Davidson2005,Davidson2006a} argue that this ``excrescent'' schwa is distinct from the both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.}

%\citet{Lees1966b,Lees1966a}

@@

@@

%\input{3.1}
%\input{3.2}
%\input{3.3}

### 2012-05-13 4.1.2.tex
include at the chapter level

If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables. For this reason, it is assumed here that affricates are single segments for the purposes of syllabification. This assumption is motivated by the fact that affricates pattern with simple segments in many languages; for instance, the complex onsets in Classical Nahua are affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.

>>>>

If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables. For this reason, an additional constraint is assumed which prevents the two components of the affricates from being split by syllabification. This is motivated by the tendency of affricates to pattern with single segments in many languages. For instance, the only complex onsets in Classical Nahua are the affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.

### 2012-05-13 4.1.3.tex
include at the chapter level

To identify the components of syllable contact clusters, the syllables identified by the above procedure must be further separated into onset, nucleus, and coda, and mapped to underlying representation. This is trivial in most cases: the nucleus is easily separated from onset and coda and surface consonants are faithful to underlying representation. The treatment of the few difficult cases is described below. Ambiguous segments are assigned to onset, nucleus, or coda, and to underlying form, using a distributional heuristic.

@@

There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right, as demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}. The strongest piece of evidence for treating the velar nasal as a pure allophone is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone. \citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here. English nasal place allophony is formalized in Section \ref{cnpasection} below.
%\citealt[][62]{Halle1985a}),

@@

The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}). The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}). 
%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
There is extensive motivation for this principle. When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː]. In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}. This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear. There is also strong external support for this principle. The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor homponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}. Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.

\begin{example}
\label{VBR}
VBR's complex onsets \citep[79--80, his transcriptions]{Buchwald2005}:

\vspace{0.5\baselineskip}
\begin{tabular}{l l l}
a. & kəræb  & `crab'  \\
   & bəlid  & `bleed' \\
b. & kəwin  & `queen' \\
   & kəwoʊt & `quote' \\
c. & kut    & `cute'  \\
   & musɪk  & `music  \\ 
\end{tabular}
\end{example}

>>>>

To identify the inventory of syllable contact clusters, it is also necessary to parse syllables into onset, nucleus, and coda. In most cases this is trivial, but a few distributional heuristics for difficult cases are described below.

@@

There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right, as demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}. The strongest piece of evidence for treating the velar nasal as a pure allophone is the general absence of ?
nset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone. \citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here. English nasal place allophony is formalized in Section \ref{cnpasection} below.
%\citealt[][62]{Halle1985a}),

@@

The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}). The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}). 
%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
There is extensive motivation for this principle. When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː]. In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}. This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear. There is also strong external support for this principle. The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor} [h]\emph{omponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}. Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.

\begin{example}[VBR's complex onsets (\citealp{Buchwald2005}:79--80, his transcriptions)]
%\protect\citep[79--80, his transcriptions]{Buchwald2005}]

\begin{tabular}{l l l}
a. & kəræb  & `crab'  \\
   & bəlid  & `bleed' \\
b. & kəwin  & `queen' \\
   & kəwoʊt & `quote' \\
c. & kut    & `cute'  \\
   & musɪk  & `music  \\ 
\end{tabular}
\label{VBR}
\end{example}

### 2012-05-13 4.3.tex
include at the chapter level

\begin{example}
Tigrinya roots with identical first and second consonants \citep{Buckley1990a}: \\
\begin{tabular}{l l l l}
a. & lʌlʌw     & `scorch'                   & (< Ge'ez \emph{lʌwlʌw} `inflame')     \\
   & mʌmʌy     & `winnow'                   & (< Ge'ez \emph{mʌymʌy} `distinguish') \\
%   & mʌmʌt & `pick out loot' & (< 
b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
\end{tabular}
\end{example}

>>>>

%\begin{example}
%Tigrinya roots with identical first and second consonants \citep{Buckley1990a}:
%\begin{tabular}{l l l l}
%a. & lʌlʌw     & `scorch'                   & (< Ge'ez \emph{lʌwlʌw} `inflame')     \\
%   & mʌmʌy     & `winnow'                   & (< Ge'ez \emph{mʌymʌy} `distinguish') \\
%   & mʌmʌt & `pick out loot' & (< 
%b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
%   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
%c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
%   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
%\end{tabular}
%\end{example}

### 2012-05-13 4.tex
more writing
include at the chapter level

\chapter{Accidental and structural gaps in English syllable contact}
\label{clusters}

@@

\begin{example}
The null hypothesis: \\
Universal grammar does not countenance constraints on the contents of URs
\end{example}

@@

\begin{quotation}
\ldots the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language. \citep[][16]{Fischer-Jorgensen1952}
\end{quotation}

@@

\begin{quotation}
Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters. The material will never be complete. It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
\end{quotation}

>>>>

%\chapter{Accidental and structural gaps in English syllable contact\protect\footnotemark{*}\protect\footnotetext{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics.}}
\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics under the same name.}}
\label{clusters}

@@

\begin{example}[null hypothesis]
Universal grammar does not countenance constraints on the contents of URs
\end{example}

@@

\begin{quote}
\ldots{}the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language. \citep[][16]{Fischer-Jorgensen1952}
\end{quote}

@@

\begin{quote}
Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters. The material will never be complete. It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
\end{quote}
